### 2024-01-02 Conclusion.tex

\chapter{Conclusion}
\label{chap:conclusion}
\minitoc

My thesis showed I did very useful things and also pleases my supervisor. I also showed that I somewhat address the challenges laid out in \chapref{chap:intro}.

\begin{figure}[htb!]
\centering
%\includegraphics[width=0.60\linewidth]{figs/University_College_London_logo}
\includegraphics[width=0.60\linewidth]{figs/kcllogo}
\caption{A super caption.\label{fig:logo}}
\end{figure}

\Figref{fig:logo} shows our incredible results. We can also have interesting equations
\begin{align}
\log_x(y) &= DL_x(\Id) \log_{\Id}(x^{-1} \circ y) \label{eq:logid}\\
\exp_{x}(\bu) &= x \circ \exp_{\Id}( DL_x(\Id)^{-1} \bu ).
\label{eq:expid}
\end{align}
and reference them as \eqref{eq:expid}.

\section{A section}
%
Youpi

\subsection{A sub-section}
%
Youpi

\subsubsection{A sub-sub-section}
%
Youpi

\paragraph{A paragraph}
%
Youpi

\section{A section}
%
Youpi

\subsection{A sub-section}
%
Youpi

\subsubsection{A sub-sub-section}
%
Youpi

\paragraph{A paragraph}
%
Youpi