### 2024-02-18 List_of_Publications.tex

\section*{First-author articles in international peer-reviewed journals}
\begin{itemize}
\item \bibentry{Bahl2023}

This also forms the basis of patent application number 2309106.9 where the first author is listed as an inventor.

\item \bibentry{Bahl2023a}
% \input{PhD-Thesis-LOP_1.bib}
%Update when 1-layer published
%Add in 2-layer paper when preprint or published
\end{itemize}

\section*{First-author articles in international conferences with published proceedings}
%Add tissue modelling reference when proceedings published

\section*{Co-author articles in international peer-reviewed journals}
\begin{itemize}
\item \bibentry{MacCormac2023}

\item \bibentry{Li2021}

\item \bibentry{Xie2021}
\end{itemize}