### 2024-02-28 Acknowledgments.tex

This thesis would not have been possible without the support of so many people. Firstly, I would like to thank my supervisors, Prof Tom Vercauteren, Dr Mads Bergholt, and Dr Jonathan Shapey, for their guidance, and the research team for their helpful advice throughout this work. I would particularly like to thank Oscar, Matt, and Jonathan for taking the time to explain clinical and anatomical concepts and allowing me to enjoy the more clinical parts of my experience so much!

I would not have enjoyed the past three and half years without so many amazing friends and family. Jo, Oscar, Ben, and Ross, thank you for brightening my days in and out of the office and making me laugh in every lunch break. Thank you to Ross especially for helping me to unscrew so many things that I happened to get stuck in the lab! Thank you to my friends from school (Charlotte, Cori, Saffy, Lucy, and Tilly), from university (especially Jenna and Branwen) and beyond, who have brought so much fun to the last few years. Thank you as well to my lovely housemates (Annabel, Alice, and Grace) for putting up with my laptop whirring away in the living room!

None of my achievements would have been possible without the continued love and support of my family. Ria and Avik who have been my partners in crime for as long as I can remember but are now giving me far more wisdom than I think I ever gave them; I am so proud of you both and I hope I’ve made you proud too. To my father for always believing in me and never failing to bring me new snacks to fuel my work! Finally, to my mother (and best friend), your love and support has brought me to where I am and who I am today.