### 2024-05-16 List_of_Publications.tex

\section*{First-author articles in international peer-reviewed journals}
\begin{itemize}
\item \bibentry{Bahl2023}

This also forms the basis of patent application number 2309106.9 where the first author is listed as an inventor.

\textcolor{red}{\item \bibentry{Bahl2023a}} \annref{3} 
% \item Anisha Bahl, Silvere Segaud, Yijing Xie, Jonathan Shapey, Mads Bergholt and Tom Vercauteren. \textit{A comparative study of analytical models of diffuse reflectance in homogeneous biological tissues: Gelatin based phantoms and monte carlo experiments.} Under review.

% \item \bibentry{Bahl2024}

\item Anisha Bahl, Jonathan Shapey, Mads Bergholt and Tom Vercauteren. \textcolor{red}{(2024).} \textit{Evaluation of a two-layer skin tissue model for diffuse reflectance:
A simulated and experimental data study.} Under \textcolor{red}{review}. \annref{3}
% \item \bibentry{Bahl2023a}
% \input{PhD-Thesis-LOP_1.bib}
%Update when 1-layer published
%Add in 2-layer paper when preprint or published
\end{itemize}

\section*{First-author articles in international conferences with published proceedings}%Add tissue modelling reference via bibtex when proceedings published 
\begin{itemize}
\item \textcolor{red}{\bibentry{Bahl2024a}} \annref{3} 
% \item \bibentry{Bahl2024a}
% \item Anisha Bahl, Silvere Segaud, Yijing Xie, Jonathan Shapey, Mads Bergholt and Tom Vercauteren. \textit{A comparative study of analytical tissue reflectance models
% using simulated and experimental data.} Awaiting publication.
\end{itemize}

\section*{Co-author articles in international peer-reviewed journals}
\begin{itemize}
\item \bibentry{MacCormac2023}

\item \bibentry{Li2021}
\end{itemize}

\section*{Co-author pre-print articles}
\begin{itemize}
\item \bibentry{Xie2021}
\end{itemize}